---
title: "Project_2"
author: "Sarah Eiben, Michelle Baumgartner, Deni Hodzic"
date: '2022-06-13'
output: html_document
---

# Install the newest version of epiwraps

```{r, eval=FALSE}
BiocManager::install("ETHZ-INS/epiwraps")
```

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(epiwraps)
})
```

# Downloading the data
In this file all the data is being downloaded from ENCODE.
The data we use is CREB peaks and signal p-values (CREB1, CREB3, CREB3L1, CREB5), cofactor peaks and signal p-values (CREBBP, JUN, EP300, HDAC2) and acetylation peaks and signal p-values (H3K27ac). We saved them into different folders.

## Downloading CREB peaks
These are all IDR thresholded bed narrowpeaks, saved in a folder named peaks_creb.

```{r, eval=FALSE}
dir.create("peaks_creb")
# CREB1
download.file("https://www.encodeproject.org/files/ENCFF193LLN/@@download/ENCFF193LLN.bed.gz", dest="peaks_creb/peaks_creb1.bed.gz")
# CREB3
download.file("https://www.encodeproject.org/files/ENCFF829IEE/@@download/ENCFF829IEE.bed.gz", dest="peaks_creb/peaks_creb3.bed.gz")
# CREB3L1
download.file("https://www.encodeproject.org/files/ENCFF118GMS/@@download/ENCFF118GMS.bed.gz", dest="peaks_creb/peaks_creb3l1.bed.gz")
# CREB5
download.file("https://www.encodeproject.org/files/ENCFF875JMR/@@download/ENCFF875JMR.bed.gz", dest="peaks_creb/peaks_creb5.bed.gz")
```

## Downloading CREB signal p-value data
These are bigwig files saved in a folder named signal_pv_creb.

```{r, eval=FALSE}
dir.create("signal_pv_creb")
# CREB1
download.file("https://www.encodeproject.org/files/ENCFF014CIM/@@download/ENCFF014CIM.bigWig", dest=("signal_pv_creb/signal_pv_creb1.bw"), mode="wb")
# CREB3
download.file("https://www.encodeproject.org/files/ENCFF462EMG/@@download/ENCFF462EMG.bigWig", dest=("signal_pv_creb/signal_pv_creb3.bw"), mode="wb")
# CREB3L1
download.file("https://www.encodeproject.org/files/ENCFF638ZCD/@@download/ENCFF638ZCD.bigWig", dest=("signal_pv_creb/signal_pv_creb3l1.bw"), mode="wb")
# CREB5
download.file("https://www.encodeproject.org/files/ENCFF335JJU/@@download/ENCFF335JJU.bigWig", dest=("signal_pv_creb/Signal_pv_creb5.bw"), mode="wb")
```

## Downloading cofactor peaks
Downloading the peaks of the following cofactors: CREBBP, JUN, EP300 and HDAC2 into a folder named peaks_cofactors.
Maybe we won't need these - we downloaded them just to be sure.

```{r, eval=FALSE}
dir.create("peaks_cofactors")
# CREBBP
download.file("https://www.encodeproject.org/files/ENCFF532VPN/@@download/ENCFF532VPN.bed.gz", dest="peaks_cofactors/peaks_crebbp.bed.gz")
# JUN
download.file("https://www.encodeproject.org/files/ENCFF865UPM/@@download/ENCFF865UPM.bed.gz", dest="peaks_cofactors/peaks_jun.bed.gz")
# EP300
download.file("https://www.encodeproject.org/files/ENCFF702XPO/@@download/ENCFF702XPO.bed.gz", dest="peaks_cofactors/peaks_ep300.bed.gz")
# HDAC2
download.file("https://www.encodeproject.org/files/ENCFF150FJT/@@download/ENCFF150FJT.bed.gz", dest="peaks_cofactors//peaks_hdac2.bed.gz")
```

## Downloading cofactor signal p-value
Download of cofactor signal p-values (CREBBP, JUN, EP300 and HDAC2) in bigwig format into a folder named signal_pc_covactors.

```{r, eval=FALSE}
dir.create("signal_pv_cofactors")
# CREBBP
download.file("https://www.encodeproject.org/files/ENCFF783COJ/@@download/ENCFF783COJ.bigWig", dest="signal_pv_cofactors/signal_pv_crebbp.bw", mode="wb")
# JUN
download.file("https://www.encodeproject.org/files/ENCFF220RDW/@@download/ENCFF220RDW.bigWig", dest="signal_pv_cofactors/signal_pv_jun.bw", mode="wb")
# EP300
download.file("https://www.encodeproject.org/files/ENCFF325DSL/@@download/ENCFF325DSL.bigWig", dest="signal_pv_cofactors/signal_pv_ep300.bw", mode="wb")
# HDAC2
download.file("https://www.encodeproject.org/files/ENCFF240UFL/@@download/ENCFF240UFL.bigWig", dest="signal_pv_cofactors/signal_pv_hdac2.bw", mode="wb")
```

## Downloading H3K27ac peaks and signal p-value
Downloading the Acetylation peaks (bed narrow peaks) and p-values (bigwig) of H3K27ac into a folder named H2K27ac.

```{r, eval=FALSE}
dir.create("H3K27ac")
# H3K27ac peaks
download.file("https://www.encodeproject.org/files/ENCFF437DPT/@@download/ENCFF437DPT.bed.gz", dest="H3K27ac/h3k27ac_peaks.bed.gz")
# H3K27ac signal p-value
download.file("https://www.encodeproject.org/files/ENCFF469JMR/@@download/ENCFF469JMR.bigWig", dest="H3K27ac/h3k27ac_signal_pv.bw", mode="wb")
```
# Reading in the peak data
In this chunk we are reading in the peak data in order to use it later.

```{r, eval=FALSE}
# CREB peaks
peaks_creb1 <- import("peaks_creb/peaks_creb1.bed.gz", format="NarrowPeak")
peaks_creb3 <- import("peaks_creb/peaks_creb3.bed.gz", format="NarrowPeak")
peaks_creb3l1 <- import("peaks_creb/peaks_creb3l1.bed.gz", format="NarrowPeak")
peaks_creb5 <- import("peaks_creb/peaks_creb5.bed.gz", format="NarrowPeak")

# Cofactor peaks
peaks_crebbp <- import("peaks_cofactors/peaks_crebbp.bed.gz", format="NarrowPeak")
peaks_jun <- import("peaks_cofactors/peaks_jun.bed.gz", format="NarrowPeak")
peaks_ep300 <- import("peaks_cofactors/peaks_ep300.bed.gz", format="NarrowPeak")
peaks_hdac2 <- import("peaks_cofactors/peaks_hdac2.bed.gz", format="NarrowPeak")

# H3K27ac peaks
peaks_h3k27ac <- import("H3K27ac/h3k27ac_peaks.bed.gz", format="NarrowPeak")
```


# Reading in the signal p-value data

This step is added in order to follow our train of thoughts and process.
First we downloaded and read everything in without the mode="bw" which does not work on windows. Hence we had to redo this time-consuming step and add the mode.

```{r, eval=FALSE}
# try something pierre told me to:
download.file("https://www.encodeproject.org/files/ENCFF783COJ/@@download/ENCFF783COJ.bigWig", dest="signal_pv_cofactors/signal_pv_crebbp.bw", mode="wb")
# CREB signal p-value
pv_creb1 <- import.bw("signal_pv_cofactors/signal_pv_crebbp.bw", format="bw")
```


```{r}
sessionInfo()
```


```{r, eval=FALSE}
save(peaks_creb1, peaks_creb3, peaks_creb3l1, peaks_creb5, peaks_crebbp, peaks_jun, peaks_ep300, peaks_hdac2, peaks_h3k27ac, file="myObjects.RData")
```
