---
title: "Project_4"
author: "Sarah Eiben"
date: '2022-06-14'
output: html_document
---

# Clustering of the signals and enrichment analysis

Help for this in the code from lecture 9.

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(ggplot2)
  library(rGREAT)
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(ensembldb)
  library(bsseq)
  library(BiocParallel)
  library(edgeR)
  library(DMRcate)
})
```
## Plotting the signals

As regions we are going to be using the peaks from the different CREB family members that we downloaded in Project_2 and the tracks are the corresponding signal p-values.
We make a list of the signal p-values of the CREB family members.
We also make a list of the peak values and merge them together to create one long list of non-overlapping regions.
Then we combine these in a matrix and create a heatmap.

```{r}
# create a list with all the peaks on chromosome 1
peak_list <- GRangesList("Peaks CREB1"=peaks_creb1_chr1, "Peaks CREB3"=peaks_creb3_chr1, "Peaks CREB3L1"=peaks_creb3l1_chr1, "Peaks CREB5"=peaks_creb5_chr1, compress=TRUE)
# merge the peaks together to create one list of non-overlapping regions
merged_peaks <- reduce(unlist(peak_list))
# make a list of the signal p-values
pv_list <- list(CREB1="signal_pv_creb/signal_pv_creb1.bw", CREB3="signal_pv_creb/signal_pv_creb3.bw", CREB3L1="signal_pv_creb/signal_pv_creb3l1.bw", CREB5="signal_pv_creb/Signal_pv_creb5.bw")
# we obtain the matrix of the signal around the regions:
m_creb1 <- signal2Matrix(pv_list, merged_peaks)
```



```{r}
sessionInfo()
```