---
title: "Project_3"
author: "Sarah Eiben"
date: '2022-06-14'
output: html_document
---

# Region access

Using the functions OverlapsAny, regionOverlaps and regionUpset we analyse the access of the DNA by the different CREB family members.

First we do everything just for Chromosome 1 to get an idea, then we move on to the whole genome.

## Restricting all of the CREB family memeber peaks to just chromosome 1

```{r}
# CREB1
peaks_creb1_chr1 <- peaks_creb1[seqnames(peaks_creb1)=="chr1"]
# CREB3
peaks_creb3_chr1 <- peaks_creb3[seqnames(peaks_creb3)=="chr1"]
# CREB3L1
peaks_creb3l1_chr1 <- peaks_creb3l1[seqnames(peaks_creb3l1)=="chr1"]
# CREB5
peaks_creb5_chr1 <- peaks_creb5[seqnames(peaks_creb5)=="chr1"]
```

## Getting an idea of the peaks in chromosome 1

Should we do this??

Using just the strongest peaks.

```{r}
#look at how many peaks there are
length(peaks_creb1_chr1)
head(peaks_creb1_chr1)
#My peaks for some reason doesn't have the qValue?
strong_peaks_creb1 <- peaks_creb1_chr1[peaks_creb1_chr1$score>median(peaks_creb1_chr1$score)]
#extend shows you a little more area around the peak, the resolution is 20, specify the regions you want to analyse
peakSignals_creb1_chr1 <- signal2Matrix("signal_pv_creb/signal_pv_creb1", regions=strong_peaks_creb1, extend=2000, w=20)
class(peakSignals_creb1_chr1[[1]])
#the data itself is a big matrix, let's look at it
peakSignals_creb1_chr1[[1]][1:5, 1:5]
plotEnrichedHeatmaps(peakSignals_creb1_chr1, raster_by_magick=FALSE)
```

## Creating a regionOverlap and a regionUpset for all CREB family members for chromosome 1

Look at lecture 5 code for this!

```{r}
listofRegions <- list(creb1=peaks_creb1_chr1, creb3=peaks_creb3_chr1, creb3l1=peaks_creb3l1_chr1, creb5=peaks_creb5_chr1)
regionOverlaps(listofRegions)
regionUpset(listofRegions)
```

## Restricting to CREB1 sites

```{r}
regionUpset(listofRegions, reference = peaks_creb1_chr1)
```

## Restricting to CREB3 sites

```{r}
regionUpset(listofRegions, reference = peaks_creb3_chr1)
```

## Restricting to CREB3L1 sites

```{r}
regionUpset(listofRegions, reference = peaks_creb3l1_chr1)
```
## Restricting to CREB5 sites

```{r}
regionUpset(listofRegions, reference = peaks_creb5_chr1)
```

# How do these plots make sense?

Explain here how we can make sense of these.
